Calendar-based graphics for visualising people's daily schedules

### 2017-07-03 main.Rmd
added some to construction and scales

Calendar-based visualisation proves useful in unfolding human-related activities in temporal context. For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days. It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office. However, their technique is too constrained to colour-encoding graphics. We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types. The proposed algorithm has been implemented into the `sugrrants` package [@sugrrants] using R [@r].

@@

## Calendar grids

@@

The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year.

@@

@@

# Applications

>>>>

Calendar-based visualisation proves useful in unfolding human-related activities in temporal context. For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days. It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office. However, their technique is too constrained to colour-encoding graphics. We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types. The proposed algorithm has been implemented in the `sugrrants` package [@sugrrants] using R [@r].

@@

## Grids setup

@@

The algorithm can be extended relatively easily to layout from one single month to multiple years depending on the time span to be visualised. The month-by-month arrangement can be determined by the user's choice through $M$ and $N$. If one would like to compare and contrast January, February and so on across multiple years, $M = 12$ and $N$ depending on the number of years could be setup.

Monthly calendar format provides an effective layout to emphasise differences between weekdays and weekends as well as day of year (for example, public holiday) simultaneously, however, either one of which is sufficient in certain visualisation applications. If there is apparent depiction of weekday effects but none of yearly effects, weekly calendar format (weeks of a year) could be considered employing; otherwise days of a month could be laid out for standing-out yearly behaviours.

(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE). Weekly and daily calendar formats are the simplifications of monthly calendar.

We only illustrate that calendar grids are laid out on the horizontal direction. The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above. It is useful for users in some countries where they get used to calendars of vertical organisation and other possible comparisons like Mondays across the columns rather than rows.

@@

One of the advantages using glyphs over heatmaps is evident in easily enabling scales on different temporal scales. All of the line glyphs shown above are drawn on the global scale so that it makes the magnitudes comparable over all the period. On the other hand, some sub-series of small magnitudes yet worth-noting shapes possibly become invisible whilst embedded in the overall picture. The individual scale or something else, as a result, helps to complement the calendar-based visualisation in general.

@@

# Grammar of graphics
